Epistemic Semantics in Guarded String Models

### 2020-06-16 tex/draft.tex
Started on bibliography.

\section{Introduction and Related Work}
Linguistic semantics in the Montague tradition proceeds by assigning
propositional {\em semantic values} to disambiguated sentences of a
natural language.  A proposition is a set or class of {\em possible
  worlds}.  These worlds are often assumed to be things with the same
nature and complexity as the world we occupy (Lewis 1986).  But
alternatively, one can work with small idealized models, in order to
illustrate and test ideas.  To build such models, spaces of worlds and
individuals are stipulated as small finite sets, and semantic values
of lexical items are constructed as functions or relations from these
small sets.  Such toy or idealized models are useful in research and
in teaching, in that it is possible to represent propositions finitely
and explicitly, and to calculate with them. The point of this paper is
to scale up toy or idealized models to countable sets of worlds, and
to constructive and computable modeling of epistemic alternatives for
agents.  We describe a certain systematic way of defining such models,
and illustrate how to apply them in natural language semantics.
The focus on epistemic semantics and clausal embedding. The fundamental move is to
identify possible worlds with strings primitive events, so that
propositions are sets of strings. An advantage in this is that it
allows for a mathematical description of an algebra of a propositions,
coupled with a computational representation using either lazy lists of
strings, or finite state machines that describe sets of strings.

@@

\begin{figure}  
\begin{tabular}[t]{l|c|c}
Amy & $e$ & ${R}_a$\\
Bob & $e$ & ${R}_b$\\
it & $d$ & $d$ \\
heads & $d \backslash_D t$ & $\lambda x.0^c{\cdot}h$ \\
tails & $d \backslash_D t$ & $\lambda x.0^c{\cdot}!h$ \\
is & $(d \backslash t)\slash(d \backslash_D t)$ & $\lambda P.\lambda x.P x$ \\
knows & $(e \backslash t)\slash_M t$ &  $\lambda p.\lambda R. {\Diamond}R p$\\
that  & \multicolumn{2}{l}{$(((e \backslash t)\slash_M t) \backslash (e \backslash t)) \slash t$} \\
& \multicolumn{2}{l}{$\lambda p.\lambda m.\lambda R.{\sim}(m ({\sim}p) R)$} \\
whether  & \multicolumn{2}{l}{$(((e \backslash t)\slash_M t) \backslash (e \backslash t)) \slash t$} \\
& \multicolumn{2}{l}{$\lambda p.\lambda m.\lambda R.{\sim}(m ({\sim}p) R)$} \\
& \multicolumn{2}{r}{$+ {\sim}(m p R)$} \\
\end{tabular}
\caption{Partial categorial grammar lexicon. The first column has a word form.
  the second column a categorial type, and third column a semantic translation in
a notation that extends the Epik term language with lambds.}
\end{figure}

>>>>

\section{Introduction and Related Work}
Linguistic semantics in the Montague tradition proceeds by assigning
propositional {\em semantic values} to disambiguated sentences of a
natural language.  A proposition is a set or class of {\em possible
  worlds}.  These worlds are often assumed to be things with the same
nature and complexity as the world we occupy \citep{lewis1986plurality}.  But
alternatively, one can work with small idealized models, in order to
illustrate and test ideas.  To build such models, spaces of worlds and
individuals are stipulated as small finite sets, and semantic values
of lexical items are constructed as functions or relations from these
small sets.  Such toy or idealized models are useful in research and
in teaching, in that it is possible to represent propositions finitely
and explicitly, and to calculate with them. The point of this paper is
to scale up toy or idealized models to countable sets of worlds, and
to constructive and computable modeling of epistemic alternatives for
agents.  We describe a certain systematic way of defining such models,
and illustrate how to apply them in natural language semantics.
The focus on epistemic semantics and clausal embedding. The fundamental move is to
identify possible worlds with strings primitive events, so that
propositions are sets of strings. An advantage in this is that it
allows for a mathematical description of an algebra of a propositions,
coupled with a computational representation using either lazy lists of
strings, or finite state machines that describe sets of strings.

@@

\begin{figure}  
\begin{tabular}[t]{l|c|c}
Amy & $e$ & ${R}_a$\\
Bob & $e$ & ${R}_b$\\
it & $d$ & $d$ \\
heads & $d \backslash_D t$ & $\lambda x.0^c{\cdot}h$ \\
tails & $d \backslash_D t$ & $\lambda x.0^c{\cdot}!h$ \\
is & $(d \backslash t)\slash(d \backslash_D t)$ & $\lambda P.\lambda x.P x$ \\
knows & $(e \backslash t)\slash_M t$ &  $\lambda p.\lambda R. {\Diamond}R p$\\
that  & \multicolumn{2}{l}{$(((e \backslash t)\slash_M t) \backslash (e \backslash t)) \slash t$} \\
& \multicolumn{2}{l}{$\lambda p.\lambda m.\lambda R.{\sim}(m ({\sim}p) R)$} \\
whether  & \multicolumn{2}{l}{$(((e \backslash t)\slash_M t) \backslash (e \backslash t)) \slash t$} \\
& \multicolumn{2}{l}{$\lambda p.\lambda m.\lambda R.{\sim}(m ({\sim}p) R)$} \\
& \multicolumn{2}{r}{$+ {\sim}(m p R)$} \\
\end{tabular}
\caption{Partial categorial grammar lexicon. The first column has a word form.
  the second column a categorial type, and third column a semantic translation in
a logical language that extends the Epik term language with lambda.}
\end{figure}
